---
title: "Analyses"
subtitle: "Project: Applicability of open science practices to completed research projects from different disciplines and research paradigms"
author: "[blinded]" # "Jürgen Schneider [0000-0002-3772-4198](https://orcid.org/0000-0002-3772-4198)"
affiliation: "[blinded]" # "University of Tübingen"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: 'hide'
    self_contained: true
---

# Packages & Setup

```{r setup, warning=F, message=F, error=F}
knitr::opts_chunk$set(
  warning = F, # don't show warnings during document generation
  message = F, # don't show messages during document generation
  error = F, # don't show errors during document generation
  echo = TRUE  # show R code
)

library(tidyverse)
library(formr)
library(viridis)
library(tidyLPA)
library(nnet)
library(kableExtra)
library(psych)
library(generics)
library(emmeans)
library(xfun)
library(here)
library(naniar)
library(stargazer)
library(ggpubr)

installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
R.Version()
```

# Data import

```{r get and join data}
formr_connect(keyring = "formr_juergen") #connect to formr

osc <- formr_raw_results(survey_name = 'osc') %>%  # import first survey
  dplyr::filter(!str_detect(session, 'XXX') & !is.na(session) & !is.na(ended)) # filter out dummy test runs

osc <- osc[-c(1:12),] # delete live test runs


# match data from discipline to data set
osc_0 <- formr_raw_results(survey_name = 'osc_0') %>% #import data from second survey
  dplyr::select(session, disc) %>%
  dplyr::mutate(disc = str_sub(disc, 1, 3))
osc <- left_join(osc, osc_0, by = "session")
```


You can get the [data on zenodo](https://doi.org/10.5281/zenodo.6834569) (not blinded) or download the data file [embeded in this HTML]().  
There is also a [data paper](https://doi.org/10.12688/f1000research.111383.2) (not blinded) available.

# Analysis of missingness

```{r}
osc_miss <- osc %>%
  dplyr::select(osp_prp:disc, -finished)

vis_miss(osc_miss)
```

All items were set to mandatory, there are no missings in the data set. 

# Variable names and value labels

Below see the variable names and their description (not variable labels), as well as value labels.  
The variable names are used throughout this document, for further details on the variables, see [codebook on zenodo](https://doi.org/10.5281/zenodo.6834569) (not blinded).

```{r}

names_labels <- data.frame(variable_name  = c(
                             "osp_cit",
                             "osp_cod",
                             "osp_dat",
                             "osp_mat",
                             "osp_met",
                             "osp_oap",
                             "osp_oer",
                             "osp_opr",
                             "osp_oso",
                             "osp_pre",
                             "osp_prp",
                             "osp_sco",
                             "osp_sof",
                             "disc",
                             "",
                             "",
                             "",
                             "",
                             "",
                             "para",
                             "",
                             "",
                             ""),
                           description =c(
                             "Citizen Science",
                             "Open Code",
                             "Open Data",
                             "Open Materials",
                             "Open Methodology",
                             "Open Access",
                             "OER",
                             "Open Peer Review",
                             "Open Source",
                             "Preregistration",
                             "Public Project Plan",
                             "Science Communication",
                             "Open Software",
                             "Discipline Cluster",
                             "",
                             "",
                             "",
                             "",
                             "",
                             "Research paradigm",
                             "",
                             "",
                             ""),
                           value_labels = c(
                             "1 = not applicable at all; 4 = highly applicable",
                             "1 = not applicable at all; 4 = highly applicable",
                             "1 = not applicable at all; 4 = highly applicable",
                             "1 = not applicable at all; 4 = highly applicable",
                             "1 = not applicable at all; 4 = highly applicable",
                             "1 = not applicable at all; 4 = highly applicable",
                             "1 = not applicable at all; 4 = highly applicable",
                             "1 = not applicable at all; 4 = highly applicable",
                             "1 = not applicable at all; 4 = highly applicable",
                             "1 = not applicable at all; 4 = highly applicable",
                             "1 = not applicable at all; 4 = highly applicable",
                             "1 = not applicable at all; 4 = highly applicable",
                             "1 = not applicable at all; 4 = highly applicable",
                             "agr = Agricultural and veterinary sciences",
                             "eng = Engineering and technology",
                             "hum = Humanities and the arts",
                             "med = Medical and health sciences",
                             "nat = Natural sciences",
                             "soc = Social sciences",
                             "mixe = explicitly mixed-methodological (equally qualitative and quantitative empirical)",
                             "qual = mainly qualitative empirical",
                             "quan = mainly quantitative empirical",
                             "none = nonempirical"
                           ))

names_labels %>%
  kbl() %>%
  kable_paper("hover")
```


# Descriptive analyses

## 13 open science practices
```{r}
descr <- describe(osc_miss[,1:13])

descr %>%
  round(3) %>%
  kbl() %>%
  kable_paper("hover")
```

## Research paradigm
```{r}
table(osc_miss$para) %>%
  round(3) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)
```

## Discipline (broad classification)
```{r}
table(osc_miss$disc) %>%
  round(3) %>%
  kbl() %>%
  kable_paper("hover", full_width = F)
```


# Latent profile analysis

## Number of profiles

LPAs based on the 13 open science practices with different number of profiles (1 to 8). Solutions are compared using BIC.

```{r}
osc %>%
  dplyr::select(osp_cit,
                osp_prp,
                osp_pre,
                osp_met,
                osp_sof,
                osp_mat,
                osp_cod,
                osp_dat,
                osp_oer,
                osp_opr,
                osp_oap,
                osp_oso,
                osp_sco) %>%
  estimate_profiles(1:8) %>%
  compare_solutions(statistics = c("BIC", "Entropy"))
```

The BIC fit index is lowest for the solutions with 4 and 5 classes. For the 5 class solution, however, the entropy falls under the threshold of .80. We therefore decide for the 4 class solution with an acceptable entropy.

## Characteristics of profiles

### Line plot of profiles

```{r, fig.dim = c(10, 6), out.width="100%"}
osc_lpa <- osc %>%
  dplyr::select(osp_oso,
                osp_sof,
                osp_cod,
                osp_pre,
                osp_oer,
                osp_cit,
                osp_prp,
                osp_opr,
                osp_sco,
                osp_mat,
                osp_dat,
                osp_oap,
                osp_met) %>%
  estimate_profiles(4) 

osc_lpa%>%
  plot_profiles(ci = NULL,
                sd = F,
                add_line = T,
                rawdata = F)
```


### Donut plots


```{r, fig.dim = c(10, 6), out.width="100%"}
# get data from LPA
lpa_data <- get_data(osc_lpa)[,c(16:20)]
lpa_data <- cbind(osc, lpa_data)
```

```{r, fig.dim = c(16, 6), out.width="100%"}
### donut plots ###############################################################

### disc ###
lpa_disc <- lpa_data %>%
  group_by(Class, disc) %>%
  dplyr::summarize(disc = disc[1],
                   disc_n = n()) %>%
  ungroup()

disc_N <- lpa_disc %>%
  group_by(Class) %>%
  summarize(N = sum(disc_n))

lpa_disc <- lpa_disc %>%
  mutate(fraction = case_when(
    Class == 1 ~ disc_n/disc_N$N[1],
    Class == 2 ~ disc_n/disc_N$N[2],
    Class == 3 ~ disc_n/disc_N$N[3],
    Class == 4 ~ disc_n/disc_N$N[4])) %>%
  group_by(Class) %>%
  mutate(ymax = cumsum(fraction),
         ymin = c(0, head(ymax, n=-1)),
         labelPosition = (ymax + ymin) / 2,
         label = paste0(disc, "\n", round(fraction * 100), "%"))



# plot for class 1
disc1 <-
ggplot(lpa_disc%>%filter(Class==1), aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=disc)) +
  geom_rect() +
  geom_text( x=2, aes(y=labelPosition, label=label, color=disc), size=6) + # x here controls label position (inner / outer)
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  coord_polar(theta="y") +
  xlim(c(-1, 4)) +
  theme_void() +
  theme(legend.position = "none") +
  ggtitle("Disciplines in class 1")


# plot for class 2
disc2 <-
ggplot(lpa_disc%>%filter(Class==2), aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=disc)) +
  geom_rect() +
  geom_text( x=2, aes(y=labelPosition, label=label, color=disc), size=6) + # x here controls label position (inner / outer)
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  coord_polar(theta="y") +
  xlim(c(-1, 4)) +
  theme_void() +
  theme(legend.position = "none")+
  ggtitle("Disciplines in class 2")

# plot for class 3
disc3 <-
ggplot(lpa_disc%>%filter(Class==3), aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=disc)) +
  geom_rect() +
  geom_text( x=2, aes(y=labelPosition, label=label, color=disc), size=6) + # x here controls label position (inner / outer)
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  coord_polar(theta="y") +
  xlim(c(-1, 4)) +
  theme_void() +
  theme(legend.position = "none")+
  ggtitle("Disciplines in class 3")

# plot for class 4
disc4 <-
ggplot(lpa_disc%>%filter(Class==4), aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=disc)) +
  geom_rect() +
  geom_text( x=2, aes(y=labelPosition, label=label, color=disc), size=6) + # x here controls label position (inner / outer)
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  coord_polar(theta="y") +
  xlim(c(-1, 4)) +
  theme_void() +
  theme(legend.position = "none")+
  ggtitle("Disciplines in class 4")


ggarrange(disc1, disc2, disc3, disc4, ncol = 4, nrow = 1)
```


```{r, fig.dim = c(16, 6), out.width="100%"}
### disc ###
lpa_para <- lpa_data %>%
  group_by(Class, para) %>%
  dplyr::summarize(para = para[1],
                   para_n = n()) %>%
  ungroup()

para_N <- lpa_para %>%
  group_by(Class) %>%
  summarize(N = sum(para_n))

lpa_para <- lpa_para %>%
  mutate(fraction = case_when(
    Class == 1 ~ para_n/para_N$N[1],
    Class == 2 ~ para_n/para_N$N[2],
    Class == 3 ~ para_n/para_N$N[3],
    Class == 4 ~ para_n/para_N$N[4])) %>%
  group_by(Class) %>%
  mutate(ymax = cumsum(fraction),
         ymin = c(0, head(ymax, n=-1)),
         labelPosition = (ymax + ymin) / 2,
         label = paste0(para, "\n", round(fraction * 100), "%"))



# plot for class 1
para1 <-
ggplot(lpa_para%>%filter(Class==1), aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=para)) +
  geom_rect() +
  geom_text( x=2, aes(y=labelPosition, label=label, color=para), size=6) + # x here controls label position (inner / outer)
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  coord_polar(theta="y") +
  xlim(c(-1, 4)) +
  theme_void() +
  theme(legend.position = "none") +
  ggtitle("Paradigms in class 1")


# plot for class 2
para2 <-
ggplot(lpa_para%>%filter(Class==2), aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=para)) +
  geom_rect() +
  geom_text( x=2, aes(y=labelPosition, label=label, color=para), size=6) + # x here controls label position (inner / outer)
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  coord_polar(theta="y") +
  xlim(c(-1, 4)) +
  theme_void() +
  theme(legend.position = "none")+
  ggtitle("Paradigms in class 2")

# plot for class 3
para3 <-
ggplot(lpa_para%>%filter(Class==3), aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=para)) +
  geom_rect() +
  geom_text( x=2, aes(y=labelPosition, label=label, color=para), size=6) + # x here controls label position (inner / outer)
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  coord_polar(theta="y") +
  xlim(c(-1, 4)) +
  theme_void() +
  theme(legend.position = "none")+
  ggtitle("Paradigms in class 3")

# plot for class 4
para4 <-
ggplot(lpa_para%>%filter(Class==4), aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=para)) +
  geom_rect() +
  geom_text( x=2, aes(y=labelPosition, label=label, color=para), size=6) + # x here controls label position (inner / outer)
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  coord_polar(theta="y") +
  xlim(c(-1, 4)) +
  theme_void() +
  theme(legend.position = "none")+
  ggtitle("Paradigms in class 4")


ggarrange(para1, para2, para3, para4, ncol = 4, nrow = 1)
```


## Predicting conditional probabilities


We predict the probability of class membership by the research paradigm and discipline of the research project. In case of significant results from a predictor, we additionally compute contrasts.

### Predicting Class 1

```{r, fig.dim = c(10, 6), out.width="100%"}
# predicting conditional probability to be in class 1
fit_predict_class1 <- lm(scale(CPROB1) ~ as.factor(para) + as.factor(disc), 
                         data = lpa_data)
summary(fit_predict_class1)
# pairwise comparisons
fit_predict_class1_em <- emmeans(fit_predict_class1, "para")  
pwpm(fit_predict_class1_em)
# fit_predict_class1_em <- emmeans(fit_predict_class1, "disc")  
# pwpm(fit_predict_class1_em)
```

### Predicting Class 2
```{r, fig.dim = c(10, 6), out.width="100%"}
# predicting conditional probability to be in class 2
fit_predict_class2 <- lm(scale(CPROB2) ~ as.factor(para) + as.factor(disc), 
                         data = lpa_data)
summary(fit_predict_class2)
# # pairwise comparisons
# fit_predict_class2_em <- emmeans(fit_predict_class2, "para")
# pwpm(fit_predict_class2_em)
# fit_predict_class2_em <- emmeans(fit_predict_class2, "disc")  
# pwpm(fit_predict_class2_em)
```

### Predicting Class 3
```{r, fig.dim = c(10, 6), out.width="100%"}
# predicting conditional probability to be in class 3
fit_predict_class3 <- lm(scale(CPROB3) ~ as.factor(para) + as.factor(disc), 
                         data = lpa_data)
summary(fit_predict_class3)
# pairwise comparisons
# fit_predict_class3_em <- emmeans(fit_predict_class3, "para")  
# pwpm(fit_predict_class3_em)
fit_predict_class3_em <- emmeans(fit_predict_class3, "disc")  
pwpm(fit_predict_class3_em)
```

### Predicting Class 4
```{r, fig.dim = c(10, 6), out.width="100%"}
# predicting conditional probability to be in class 4
fit_predict_class4 <- lm(scale(CPROB4) ~ as.factor(para) + as.factor(disc), 
                         data = lpa_data)
summary(fit_predict_class4)
# pairwise comparisons
fit_predict_class4_em <- emmeans(fit_predict_class4, "para")  
pwpm(fit_predict_class4_em)
# fit_predict_class4_em <- emmeans(fit_predict_class4, "disc")  
# pwpm(fit_predict_class4_em)
```

### Table with all 4 models

```{r, results='asis'}
stargazer(fit_predict_class1, 
          fit_predict_class2, 
          fit_predict_class3, 
          fit_predict_class4, 
          type = "html",
          report = "vcsp",
          p.auto = F,
          font.size = "small",
          intercept.bottom = F,
          dep.var.labels.include = F,
          column.labels = c("--*p[Cluster 1]*--", "--*p[Cluster 2]*--", 
                            "--*p[Cluster 3]*--", "--*p[Cluster 4]*--"),
          notes = "",
          notes.label = "",
          notes.append = F
          )
```


